Using Genetic Algorithms to Optimize the Analogue Method for Precipitation Prediction in the Swiss Alps

### 2016-12-21 Optimization case study - JoH/elsarticle-template.tex
JoH paper under correction. Removing tables and figures.

\title{Using Genetic Algorithms to Optimize the Analog Method for Precipitation Downscaling in the Swiss Alps}

@@

\begin{abstract}
Analog methods aim at predicting precipitation based on synoptic predictors provided by general circulation models or numerical weather prediction models. The method samples a selection of days in the archives that are similar to the target day to predict, and consider the set of their corresponding observed precipitation (the predictand) as the conditional distribution for the target day. The relationship between predictors and predictands relies on some parameters that characterize how and where the similitude between two atmos+pheric situations is defined.

This relationship is usually established by a semi-automatic sequential procedure that has strong limitations: (i) it cannot automatically choose the pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom. A new global optimization approach relying on genetic algorithms can optimize all parameters jointly and automatically, which is a breakthrough in the way the analog method was calibrated until now. It allows taking into account parameters inter-dependencies, and selecting objectively some parameters that were manually selected beforehand (such as the pressure levels and the temporal windows of the predictor variables).

@@

\begin{keyword}
Precipitation prediction\sep
Precipitation downscaling\sep
Analog method\sep
Optimization\sep
Genetic algorithms\sep
Alpine climate
\end{keyword}

@@

The analog method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969}, that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015, BenDaoud2016}, or a climate context \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

In operational forecasting, it has been used mainly by practitioners, notably hydropower companies or flood forecasting services, while in climate studies, it is used to downscale the results of global climate model (GCM) simulation runs. In this last case, although the GCM represents the large or regional scale evolutions of the atmosphere, it would be far too computer intensive to resolve variables such as precipitation at a rather small and representative scale. Therefore the idea is to rely on past observations to bypass these unaffordable small scale simulations and to go from the large scale situation proposed by the GCM to the end variables like precipitation by searching analogs in the archives.

The method can be designed with multiple successive subsampling steps, each of them relying on different meteorological variables. A certain number of parameters define the relationship between predictors and predictands, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criterion itself, and finally the number of analog situations to keep at each subsampling level. These parameters are usually calibrated by means of a semi-automatic sequential procedure \citep[see][for the details]{Bontron2004, Radanovics2013, Horton2016a}, i.e. by optimizing each single parameter one at a time, in an arbitrarily chosen order, with no or little reassessment. This sequential approach has therefore strong limitations: (i) it cannot automatically choose the optimal pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom, such as a weighting between the pressure levels. Thus, even if the processing involved is relatively fast, the sequential approach requires laborious assessments of predictors combinations (variables, pressure levels, temporal windows), and presents a high risk of ending in a local optimum due to subjective initial choices and lack of consideration of parameter inter-dependencies. Other calibration methods exist for specific applications, such as the radar images \citep{Panziera2011, Foresti2015a}.

With the perspective to overcome these limitations, a global optimization by genetic algorithms (GAs) was introduced by \citet{Horton2016}. An intensive assessment work resulted in recommendations of parametrization of GAs in order to optimize AMs successfully. The present paper is based on these recommendations and illustrate them for precipitation prediction for the upper Rh\^{o}ne catchment in the Swiss Alps. It aims at proving the relevance of a fully-automatic, objective, and global, optimization technique for AMs. The applications are indeed numerous, as the AM has to be adapted to every new location it is applied, or to any new predictand it should predict.

@@

\subsection{The analog method}
\label{sec:references}

@@

The $N_{1}$ dates with the lowest values of S1 are considered as analogs to the target day. The number of analogs, $N_{1}$, is a parameter to calibrate. Then, the daily observed precipitation amount of the $N_{1}$ resulting dates provide the empirical conditional distribution considered as the probabilistic prediction for the target day.

The other most know parametrization adds a second level of analogy on moisture variables (method 2Z-2MI, Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a moisture index made of the product of the precipitable water with the relative humidity at 850~hPa (RH850). \cite{Horton2012a} confirmed that this index is also better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis considered independently. When adding a second level of analogy, $N_{2}$ dates are subsampled in the $N_{1}$ analogs on the atmospheric circulation, to end up with a smaller number of analog situations. When a second level of analogy is added, a higher number of analogs $N_{1}$ is kept on the first level. Moisture fields are not as well predicted by NWP models as pressure variables. It means that the 2Z-2MI method, when used in real-time forecasting, is very dependent on the skill of the NWP model in predicting moisture fields, and thus is often used on limited lead times.

@@

The performance assessment in the present context consists in verifying the prediction of an ensemble-probabilistic technique. The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $n$ can be written:

@@

The reference method for the analogy of the atmospheric circulation (2Z, Table \ref{table:params_R1}), based on Z500 and Z1000, was first considered. The optimizer had to choose simulatneously the number of analogs, both spatial windows with no overlapping constraint (they can differ from one pressure level to another), as well as the temporal windows (hours of observation of the geopotential), which cannot be achieved with the sequential calibration technique. With these degrees of freedom, a relative CRPSS improvement of 3.97\% and 2.45\% in average was obtained for the CP and the VP respectively, with regard to the reference method based on Z500 and Z1000. Some tests showed that most of the gains are due to the non-overlapping spatial windows. This is not a large improvement, but it demonstrated that the optimizer was able to get relevant parameters.

Then, an additional degree of freedom was provided to the GAs by letting them choose the pressure levels along with the other parameters (analog numbers, spatial and temporal windows), which is also a non-automated process in the sequential calibration. This degree of freedom increased the optimization time and may decrease the number of simulations that converge to a single solution. However, most solutions were very close in terms of score. The averaged relative improvement of the CRPSS is 5.63\% for the CP and 3.82\% for the VP. The pressure levels that were chosen are Z500 or Z700 for the upper level, and Z925 or Z1000 (most often) for the lower level.

@@

\begin{itemize}
	\setlength\itemsep{-4px}
	\item the selection of the pressure levels (4 degrees)
	\item the temporal windows (4 degrees)
	\item the spatial windows (4x4 degrees)
	\item the weights (4 degrees)
	\item the number of analogs (1 degree).
\end{itemize}

@@

The resulting optimized parameters for 4Zo vary from one subregion to another. An example of the detailed parameters is provided for the Swiss Chablais in Table \ref{table:params_GA_4Zo}. The optimized spatial windows are given for every subregion in Figure \ref{fig:spatial_windows_4Zo}, and the selected pressure levels in Table~\ref{table:levels_GA_4Zo}.

@@

The CRPS score was also discretized into its accuracy and sharpness components, as suggested by \citet{Bontron2004}. The changes in each of these components has been illustrated in Figure \ref{fig:figure_dcrps_comp_4Zo} relatively to the total CRPS value of the 2Z reference method for the different regions and both the CP and the VP. One can see that the accuracy part is always improved to a greater extent than the sharpness, which can occasionally deteriorate. It means that the medians of the predicted precipitations distributions are closer to the observed value, whereas the spread of the distribution vary, but is in general a bit narrower.

The analysis of the parameters cross-compatibility shows that obviously, the parameters are optimal on the CP when they are optimized for a given subregion (Figure \ref{fig:crossing_4Zo_calib}). However, the losses in CRPSS when exchanging the parameters are not of the same magnitude between the different subregions. Indeed, the Upper Rhone Valley (7) and moreover the Southeast ridges (8) seem to behave significantly differently, likely due to their particular sensitivity to southerly flows \citep{Horton2012}. This points at the importance of taking into account leading meteorological influences during discretization, that are not always best represented by the physical distance. Globally, the same pattern can be observed for the VP (Figure \ref{fig:crossing_4Zo_valid}), but in this case, minor improvements may occur when crossing the parameters, due to the presence of other events in the VP that may be better predicted by a different parameter set. The relatively small gaps in score between parameterizations indicate that even though the parameters may differ significantly, the performance may not be drastically affected. Even a change in the pressure level does not mean a radical drop of the score value. A different parametrization may lead to a distinct selection of analog days, and thus to an improvement of the prediction under certain weather conditions at the expense of others.

@@

As previously, the optimized parameters differ from one subregion to another, and this even to a greater extent. Detailed examples are again provided for the Swiss Chablais subregion in Table \ref{table:params_GA_4Zo_2MIo} for 4Zo-2MIo. The resulting spatial windows are displayed in Figure \ref{fig:spatial_windows_4Zo-2MIo} for 4Zo-2MIo, along with the selected pressure levels for both the circulation and the moisture analogy (Table \ref{table:levels_GA_4Zo_2MIo}).

@@

When optimizing a method made of 2 levels of analogy, the introduction of moisture variables in the second level has an influence on the parameter values of the first level. It means that the two levels of analogy bring complementary information, and are thus not independent. This is first visible on the number $N_{1}$ of analogs to be selected on the first level, and on the selection of the pressure levels for the circulation analogy.

As for the sequential procedure, the optimal value of $N_{1}$ increases when adding a second level of analogy (Figure \ref{fig:figure_nb_analogs}). One can also notice that the optimal number of analogs $N_{2}$ for the second level of analogy of 4Zo-2MIo is slightly inferior to $N_{1}$ from 4Zo, but very close. There is a trend between the optimal analog number values of both methods: the higher $N_{1}$ of the 4Zo method, the higher $N_{1}$ and $N_{2}$ of 4Zo-2MIo is (Figure \ref{fig:figure_nb_analogs_relationships}). This relationship is not to be considered perfectly robust and should not be transposed to another case study, but in this case, it relates the magnitudes of the $N_{1}$ and $N_{2}$ on the various levels and methods.

The optimal final number of analogs do not vary much: $23 \leq N_{1} \leq 33$ for 4Zo and $21 \leq N_{2} \leq 28$ for 4Zo-2MIo. However, the optimal number of $N_{1}$ analogs of the first level of 4Zo-2MIo varies to a greater extent: $48 \leq N_{1} \leq 84$. In this latter method, it may be problematic to consider a fixed and unique value for all regions.

@@

As previously, the CRPS scores were discretized into their sharpness and accuracy components. The changes in each of these components, expressed relatively to the total CRPS values, is once again in favor of the accuracy over the sharpness (Figure \ref{fig:figure_dcrps_comp_4Zo-2HIo} for 4Zo-2MIo; 4Zo-4MIo not shown, but similar).

The analysis of the parameters cross-compatibility (Figures \ref{fig:crossing_4Zo-2MIo_calib} and \ref{fig:crossing_4Zo-2MIo_valid}) is also very similar to the one on the circulation analogy only (results for 4Zo-4MIo not shown as very similar). The same pattern can be observed, with a drop of performance for the subregions submitted to different meteorological influences. However, the losses of performance are globally more important than before, suggesting that more complex methods with moisture variables are less transposable to another subregion, even though both are located within the same grid cell of the reanalysis dataset.

@@

\begin{itemize}
	\item Using 4 pressure levels for the analogy of circulation. It seems to be the optimal number for the studied region. Beyond that value the validation score drops, revealing a loss in robustness due to over-parametrization.
	\item The automatic and joint optimization of all parameters: the analogs number, the selection of the pressure levels and the temporal windows, and the spatial windows. These parameters are highly interdependent, so one needs to optimize them jointly in order to identify optimal combinations. Traditional calibration procedures based on a systematic assessment of every combination is not possible anymore when considering more than 2 pressure levels.
	\item The introduction of distinct spatial windows between pressure levels. The synoptic circulation is characterized by features with very different scales depending on the height, and important information for predicting rainfall is not necessarily located in the same area from one level to another. However, the optimized spatial windows are consistent in between the subregions.
	\item The weighting of the analogy criteria between different pressure levels. It can be influenced by the variability of the geopotential with altitude, and the change of some levels significance with the targeted region. There is a trend for the weighting of circulation predictors to decrease with the increase in pressure, as one can see in Figure \ref{fig:levels_weights} for the method 4Zo, and in Figure \ref{fig:levels_weights_average} for the averages over the 3 methods. However, the values stay around equity. This may not be the most influencing factor, and we may suggest to remove it first when trying to reduce the number of degrees of freedom.
	\item The joint optimization of the circulation and moisture analogy levels, that are usually calibrated successively. We have been able to demonstrate that there is a dependency between the analogy levels, and that in order to approach the optimal parameters, one must consider them jointly.
\end{itemize}

GAs have proved very useful to optimize complex variants of the AM, and to assess new degrees of freedom that were not available so far. However, it can be dangerous to add too many parameters to optimize. Indeed, the optimizer will probably use them to successfully improve the calibration score, but the validation control remains very important in order to determine if one is actually improving the method, or if it is being over-parametrized.

@@

We tried to optimize the preselection period (i.e. the 4-months window) jointly with the other parameters, but no improvement was observed. Optimizing the moisture flux, which is composed of the moisture index multiplied with the wind flux, was also assessed. However, the results were not better than when considering the moisture index alone. This may be related to the fact that the optimizer tries to provide the best analogy of the atmospheric circulation in the first place, which makes the wind information less relevant in the second level of analogy.

@@

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig04.pdf}}
	\caption{Changes in the CRPS components of 4Zo relatively to the total CRPS of the reference methods 2Z (Table \ref{table:params_R1}) for the CP and the VP.}
	\label{fig:figure_dcrps_comp_4Zo}
\end{figure}

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig10.pdf}}
	\caption{Changes in the CRPS components of 4Zo-2MIo relatively to the total CRPS of the reference methods 2Z-2MI (Table \ref{table:params_R2}) for the CP and the VP.}
	\label{fig:figure_dcrps_comp_4Zo-2HIo}
\end{figure}

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig11.pdf}}
	\caption{Number of analogs for the different regions and the various methods resulting from the optimization. Method 4Zo is made of a single level of analogy with $N_{1}$ analogs, whereas 4Zo-2MIo has two levels of analogy with respectively $N_{1}$ and $N_{2}$ analogs.}
	\label{fig:figure_nb_analogs}
\end{figure}

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig12.pdf}}
	\caption{Relationships between the optimal number of analogs for both levels of the 4Zo-2MIo method and the corresponding ones for the unique level of the 4Zo method.}
	\label{fig:figure_nb_analogs_relationships}
\end{figure}

@@

\begin{table}[t]
	\caption{Parameters of the reference method on the atmospheric circulation (2Z). The first column is the level of analogy (0 for preselection), then comes the meteorological variable and its hour of observation within the target day (temporal window). The criterion used for the current level of analogy is then provided, as well as the number of analogs.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline
			Level & Variable & Hour & Criteria & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{50} \\
			& Z500 & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R1}
\end{table}

@@

\begin{table}[t]
	\caption{Parameters of the 4Zo method (analogy on 4 levels of the atmospheric circulation) optimized for the Chablais subregion. The columns are the following: L = level of analogy, V = meteorological variable, H = hour of observation within the target day, D = domain or spatial window, W = weighting of the score for the selected pressure levels, C = criterion, N = number of analogs.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline L & V & H & D & W & C & N \\ 
			\hline 
			0 & \multicolumn{6}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{8}{*}{1} &  \multirow{2}{*}{Z300} & \multirow{2}{*}{12~h} & -2.5 - 15.0 \degree E & \multirow{2}{*}{32\%} & \multirow{8}{*}{S1} & \multirow{8}{*}{28} \\
			& & & 37.5 - 57.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z700} & \multirow{2}{*}{24~h} & 2.5 - 10.0 \degree E & \multirow{2}{*}{22\%} & & \\ 
			& & & 42.5 - 45.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{6~h} & -5.0 - 15.0 \degree E & \multirow{2}{*}{24\%} & & \\ 
			& & & 42.5 - 47.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{30~h} & 2.5 - 15.0 \degree E & \multirow{2}{*}{22\%} & & \\ 
			& & & 40.0 - 50.0 \degree N & & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_GA_4Zo}
\end{table}

>>>>

\title{Using Genetic Algorithms to Optimize the Analogue Method for Precipitation Prediction in the Swiss Alps}

@@

\begin{abstract}
Analogue methods aim at predicting precipitation based on synoptic predictors provided by general circulation models or numerical weather prediction models. The method samples a selection of days in the archives that are similar to the target day to predict, and consider the set of their corresponding observed precipitation (the predictand) as the conditional distribution for the target day. The relationship between predictors and predictands relies on some parameters that characterize how and where the similitude between two atmos+pheric situations is defined.

This relationship is usually established by a semi-automatic sequential procedure that has strong limitations: (i) it cannot automatically choose the pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom. A new global optimization approach relying on genetic algorithms can optimize all parameters jointly and automatically, which is a breakthrough in the way the analogue method was calibrated until now. It allows taking into account parameters inter-dependencies, and selecting objectively some parameters that were manually selected beforehand (such as the pressure levels and the temporal windows of the predictor variables).

@@

\begin{keyword}
Precipitation prediction\sep
Precipitation downscaling\sep
Analogue method\sep
Optimization\sep
Genetic algorithms\sep
Alpine climate
\end{keyword}

@@

The analogue method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969}, that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015, BenDaoud2016}, or a climate context \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

In operational forecasting, it has been used mainly by practitioners, notably hydropower companies or flood forecasting services, while in climate studies, it is used to downscale the outputs of general circulation model (GCM) or regional climate models (RCMs) simulation runs \citep{Dayon2015} or to reconstruct past weather conditions \citep{Caillouet2016}. In future climate studies, RCMs are often used to dynamically downscale precipitation to a local scale. However, even though the relevance of RCMs'outputs increases, precipitation processes are parametrized and a bias correction of the outputs is still often required. Moreover, their application is computer intensive and cannot cover the full range of GCMs products. Therefore the idea is to rely on past observations to bypass the small scale simulations and to go from the large scale situation to the end variables like precipitation by searching analogues in the archives.

A big advantage of AMs is that they create realistic precipitation patterns for a region, provided that the analogue dates are the same, since they are based on observed situations with consistent spatial distribution \citep{Radanovics2013, Chardon2014}. For the same reason, they can also provide multivariate predictions that are physically consistent \citep{Raynaud2016b}.

The method can be designed with multiple successive subsampling steps, each of them relying on different meteorological variables. A certain number of parameters define the relationship between predictors and predictands, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criterion itself, and finally the number of analogue situations to keep at each subsampling level. These parameters are usually calibrated by means of a semi-automatic sequential procedure \citep[see][for the details]{Bontron2004, Radanovics2013, Horton2016a}, i.e. by optimizing each single parameter one at a time, in an arbitrarily chosen order, with no or little reassessment. This sequential approach has therefore strong limitations: (i) it cannot automatically choose the optimal pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom, such as a weighting between the pressure levels. Thus, even if the processing involved is relatively fast, the sequential approach requires laborious assessments of predictors combinations (variables, pressure levels, temporal windows), and presents a high risk of ending in a local optimum due to subjective initial choices and lack of consideration of parameter inter-dependencies. Other calibration methods exist for specific applications, such as the radar images \citep{Panziera2011, Foresti2015a}.

With the perspective to overcome these limitations, a global optimization by genetic algorithms (GAs) was introduced. An intensive assessment work resulted in recommendations of parametrization of GAs in order to optimize AMs successfully \citet{Horton2016}. The present paper is based on these recommendations and apply them for precipitation prediction for the upper Rh\^{o}ne catchment in the Swiss Alps. It aims at proving the relevance of a fully-automatic, objective, and global, optimization technique for AMs. The applications are indeed numerous, as the AM has to be adapted to every new location it is applied, or to any new predictand it should predict.

@@

\subsection{The analogue method}
\label{sec:references}

@@

The $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day. The number of analogues, $N_{1}$, is a parameter to calibrate. Then, the daily observed precipitation amount of the $N_{1}$ resulting dates provide the empirical conditional distribution considered as the probabilistic prediction for the target day.

The other most know parametrization adds a second level of analogy on moisture variables (method 2Z-2MI, Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a moisture index made of the product of the precipitable water with the relative humidity at 850~hPa (RH850). \cite{Horton2012a} confirmed that this index is also better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis considered independently. When adding a second level of analogy, $N_{2}$ dates are subsampled in the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When a second level of analogy is added, a higher number of analogues $N_{1}$ is kept on the first level. Moisture fields are not as well predicted by NWP models as pressure variables. It means that the 2Z-2MI method, when used in real-time forecasting, is very dependent on the skill of the NWP model in predicting moisture fields, and thus is often used on limited lead times.

@@

The performance assessment in the present context consists in verifying the prediction of an ensemble-probabilistic technique. The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $n$ can be written:

@@

The reference method for the analogy of the atmospheric circulation (2Z, Table \ref{table:params_R1}), based on Z500 and Z1000, was first considered. The optimizer had to choose simulatneously the number of analogues, both spatial windows with no overlapping constraint (they can differ from one pressure level to another), as well as the temporal windows (hours of observation of the geopotential), which cannot be achieved with the sequential calibration technique. With these degrees of freedom, a relative CRPSS improvement of 3.97\% and 2.45\% in average was obtained for the CP and the VP respectively, with regard to the reference method based on Z500 and Z1000. Some tests showed that most of the gains are due to the non-overlapping spatial windows. This is not a large improvement, but it demonstrated that the optimizer was able to get relevant parameters.

Then, an additional degree of freedom was provided to the GAs by letting them choose the pressure levels along with the other parameters (analogue numbers, spatial and temporal windows), which is also a non-automated process in the sequential calibration. This degree of freedom increased the optimization time and may decrease the number of simulations that converge to a single solution. However, most solutions were very close in terms of score. The averaged relative improvement of the CRPSS is 5.63\% for the CP and 3.82\% for the VP. The pressure levels that were chosen are Z500 or Z700 for the upper level, and Z925 or Z1000 (most often) for the lower level.

@@

\begin{itemize}
	\setlength\itemsep{-4px}
	\item the selection of the pressure levels (4 degrees)
	\item the temporal windows (4 degrees)
	\item the spatial windows (4x4 degrees)
	\item the weights (4 degrees)
	\item the number of analogues (1 degree).
\end{itemize}

@@

The resulting optimized parameters for 4Zo vary from one subregion to another. The optimized spatial windows are given for every subregion in Figure \ref{fig:spatial_windows_4Zo}, and the selected pressure levels in Table~\ref{table:levels_GA_4Zo}.

@@

The analysis of the parameters cross-compatibility shows that obviously, the parameters are optimal on the CP when they are optimized for a given subregion (Figure \ref{fig:crossing_4Zo_calib}). However, the losses in CRPSS when exchanging the parameters are not of the same magnitude between the different subregions. Indeed, the Upper Rhone Valley (7) and moreover the Southeast ridges (8) seem to behave significantly differently. These two regions have different climate properties than the others as they are particularly sensitive to southerly flows. Indeed, almost all heavy precipitation events occurred under a southerly regime, such as the Liguria, Piémont and Aosta regions in Italy, whereas the other subregions of the catchment had heavy precipitation events mainly under a westerly regime \citep{Horton2012}. Thus, as the performance score is significantly influenced by heavy precipitation values, the parameters for the different subregions are optimized to better predict these days. It can then be expected that the optimal parameters differ between these two subregions and the other ones. This points at the importance of taking into account leading meteorological influences during discretization, that are not always best represented by the physical distance.

Globally, the same pattern can be observed for the VP (Figure \ref{fig:crossing_4Zo_valid}), but in this case, minor improvements may occur when crossing the parameters, due to the presence of other events in the VP that may be better predicted by a different parameter set. The relatively small gaps in score between parameterizations indicate that even though the parameters may differ significantly, the performance may not be drastically affected. Even a change in the pressure level does not mean a radical drop of the score value. A different parametrization may lead to a distinct selection of analogue days, and thus to an improvement of the prediction under certain weather conditions at the expense of others.

@@

As previously, the optimized parameters differ from one subregion to another, and this even to a greater extent. The resulting spatial windows are displayed in Figure \ref{fig:spatial_windows_4Zo-2MIo} for 4Zo-2MIo, along with the selected pressure levels for both the circulation and the moisture analogy (Table \ref{table:levels_GA_4Zo_2MIo}).

@@

When optimizing a method made of 2 levels of analogy, the introduction of moisture variables in the second level has an influence on the parameter values of the first level. It means that the two levels of analogy bring complementary information, and are thus not independent. This is first visible on the number $N_{1}$ of analogues to be selected on the first level, and on the selection of the pressure levels for the circulation analogy.

As for the sequential procedure, the optimal value of $N_{1}$ increases when adding a second level of analogy (Figure \ref{fig:figure_nb_analogs}). One can also notice that the optimal number of analogues $N_{2}$ for the second level of analogy of 4Zo-2MIo is slightly inferior to $N_{1}$ from 4Zo, but very close. There is a trend between the optimal analogue number values of both methods: the higher $N_{1}$ of the 4Zo method, the higher $N_{1}$ and $N_{2}$ of 4Zo-2MIo is (Figure \ref{fig:figure_nb_analogs_relationships}). This relationship is not to be considered perfectly robust and should not be transposed to another case study, but in this case, it relates the magnitudes of the $N_{1}$ and $N_{2}$ on the various levels and methods.

The optimal final number of analogues do not vary much: $23 \leq N_{1} \leq 33$ for 4Zo and $21 \leq N_{2} \leq 28$ for 4Zo-2MIo. However, the optimal number of $N_{1}$ analogues of the first level of 4Zo-2MIo varies to a greater extent: $48 \leq N_{1} \leq 84$. In this latter method, it may be problematic to consider a fixed and unique value for all regions.

@@

The analysis of the parameters cross-compatibility (Figures \ref{fig:crossing_4Zo-2MIo_calib} and \ref{fig:crossing_4Zo-2MIo_valid}) is also very similar to the one on the circulation analogy only (results for 4Zo-4MIo not shown as very similar). The same pattern can be observed, with a drop of performance for the subregions submitted to different meteorological influences. However, the losses of performance are globally more important than before, suggesting that more complex methods with moisture variables are less transposable to another subregion, even though both are located within the same grid cell of the reanalysis dataset. Moisture fields have greater variability than pressure fields, and thus a change in the spatial windows can have a greater impact on the method performance. Indeed, the two regions with the lowest cross-compatibility are the Upper Rhone Valley (7) and the Southeast ridges (8), which have similar optimal pressure levels and temporal windows to other regions, but have rather different spatial windows on the moisture predictor.

@@

\begin{itemize}
	\item Using 4 pressure levels for the analogy of circulation. It seems to be the optimal number for the studied region. Beyond that value the validation score drops, revealing a loss in robustness due to over-parametrization.
	\item The automatic and joint optimization of all parameters: the analogues number, the selection of the pressure levels and the temporal windows, and the spatial windows. These parameters are highly interdependent, so one needs to optimize them jointly in order to identify optimal combinations. Indeed, there is a strong interdependence between space and time in the atmospheric circulation, so that e.g. the spatial window should move upstream the main atmospheric flow for earlier temporal windows.	Traditional calibration procedures based on a systematic assessment of every combination is not possible anymore when considering more than 2 pressure levels.
	\item The introduction of distinct spatial windows between pressure levels. The synoptic circulation is characterized by features with very different scales depending on the height, and important information for predicting rainfall is not necessarily located in the same area from one level to another. However, the optimized spatial windows are consistent in between the subregions.
	\item The weighting of the analogy criteria between different pressure levels. It can be influenced by the variability of the geopotential with altitude, and the change of some levels significance with the targeted region. There is a trend for the weighting of circulation predictors to decrease with the increase in pressure, as one can see in Figure \ref{fig:levels_weights} for the method 4Zo, and in Figure \ref{fig:levels_weights_average} for the averages over the 3 methods. However, the values stay around equity. This may not be the most influencing factor, and we may suggest to remove it first when trying to reduce the number of degrees of freedom.
	\item The joint optimization of the circulation and moisture analogy levels, that are usually calibrated successively. We have been able to demonstrate that there is a dependency between the analogy levels, and that in order to approach the optimal parameters, one must consider them jointly.
\end{itemize}

GAs have proved very useful to optimize complex variants of the AM, and to assess new degrees of freedom that were not available so far. However, it can be dangerous to add too many parameters to optimize. Indeed, the optimizer will probably use them to successfully improve the calibration score, but the validation control remains very important in order to determine if one is actually improving the method, or if it is being over-parametrized. Moreover, it might not always be desirable to increase the number of degrees of freedom, and some constraints (e.g. same weighting of the analogy criteria between different pressure levels) can be justified. However, one should first assess the consequence of a constraint before establishing it. In this sense, even though not all degrees of freedom can be found useful, GAs allow assessing their influence.

@@

We tried to optimize the length of the preselection period (i.e. the seasonal stratification, which was a 4-month window until now) jointly with the other parameters, but no improvement was observed. Optimizing the moisture flux, which is composed of the moisture index multiplied with the wind flux, was also assessed. However, the results were not better than when considering the moisture index alone. This may be related to the fact that the optimizer tries to provide the best analogy of the atmospheric circulation in the first place, which makes the wind information less relevant in the second level of analogy.

@@

AMs optimized with GAs show an improvement of the prediction for days with heavier precipitation, including extremes. However, even though the distribution of analogue precipitation values moved towards the targeted extreme, no new extreme values were added to the sample of candidate situations. The extremes in AMs can be modelled by, for example, extrapolation of a truncated exponential or gamma distribution fitted to the analogue values \citep{Obled2002}. Another possible approach is by combining AMs with other methods \citep[e.g.][]{Chardon2014}.

@@

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig11.pdf}}
	\caption{Number of analogues for the different regions and the various methods resulting from the optimization. Method 4Zo is made of a single level of analogy with $N_{1}$ analogues, whereas 4Zo-2MIo has two levels of analogy with respectively $N_{1}$ and $N_{2}$ analogues.}
	\label{fig:figure_nb_analogs}
\end{figure}

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig12.pdf}}
	\caption{Relationships between the optimal number of analogues for both levels of the 4Zo-2MIo method and the corresponding ones for the unique level of the 4Zo method.}
	\label{fig:figure_nb_analogs_relationships}
\end{figure}

@@

\begin{table}[t]
	\caption{Parameters of the reference method on the atmospheric circulation (2Z). The first column is the level of analogy (0 for preselection), then comes the meteorological variable and its hour of observation within the target day (temporal window). The criterion used for the current level of analogy is then provided, as well as the number of analogues.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline
			Level & Variable & Hour & Criteria & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{50} \\
			& Z500 & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R1}
\end{table}

@@
